Tail Indices and Scale Parameters in Financial Time Series

### 2017-02-21 papers/FX/draft.tex
Started the section of power-utility functions with $\gamma > 1$.
progress on the monotonicity proof.

\section{Generalized disappointment aversion}
\label{sec:GDA}
Routledge and Zin (2010) \cite{routledge2010generalized} define the GDA
preferences over discrete states in their formula (2). A slight extension
is to consider a continuum of possible states. Suppose the state variable
$C$ is continuously supported on $(0, \infty)$ and has distribution
function $G(\cdot)$. The utility function of an agent exhibiting GDA
preferences reads
\begin{equation}
  u(v)= \E u(C) - b \int_{0}^{\delta v}
  \left[ u(\delta v) - u(x) \right] dG(x)\label{11}%
\end{equation}
where $\delta > 0$, and $v>0$. Equivalently
\begin{equation}
  \label{eq:xxie0}
  u(v) = \E u(C) - b \E[u(\delta v) - u(C); C < \delta v]
\end{equation}
Here $v$ can be thought of as the certainty payoff equivalent to the
risky payoff $C$. Note that when $b=0$, 
preferences are expected utility; when $\delta=1$ and $b>0$,
preferences follow Gul's (1991) original disappointment aversion.

\subsection{power-law utility function}
In this section we follow Routledge and Zin \cite{routledge2010generalized} and assume
that the utility function of consumption $C$ is a power function%
\begin{equation}
  \label{eq:power_utility}
  u(C)=\frac{1}{1-\gamma}C^{1-\gamma}%  
\end{equation}
where $\gamma > 0$. For simplicity we assume the investor
initially has one unit of wealth. Let $r$ be the return on the
riskless state bond and $X$ be the return on equity. Portfolio weights
are $\left(1-\phi\right)$ and $\phi$ respectively. Then
\begin{equation*}
  %% u(C)=\frac{1}{1-\gamma}V^{1-\gamma} \max \left\{
  %% 0, 1 + (1-\phi) r  +\phi X
  %% \right\}^{1-\gamma} \label{3}%
  C = (1 - \phi) e^r + \phi e^X
\end{equation*}

\subsubsection{mildly risk-avert utility functions}
We first consider the class of utility functions for which $\gamma < 1$.
In this case $\E u(C) < \infty$ only if the distribution of $X$, call
it $F(\cdot)$, has a light upper tail. We assume this for the rest of
this section.

We note that utility functions with $\gamma < 1$ represent investors
who are less risk-avert than are those represented by utility
functions with $\gamma > 1$.

Formally, we assume
\begin{equation*}
  F(x) = \left\{
  \begin{array}{lr}
    p M^\alpha/|x|^\alpha & \text{ if } x < -M \\
    F_u(x) & \text{ if } x \geq -M \\
  \end{array}
  \right.
\end{equation*}
where $0 < p < 1$ is a constant; $\alpha > 1$, $M > 1$ and $F_u(\cdot)$
is some distribution function satisfying
\begin{itemize}
\item $F_u(-M) = p$
\item
  \[
  \int_{-M}^\infty [(1 - \phi) e^r + \phi e^x]^{1 - \gamma} d F_u(x) < \infty
  \]
\end{itemize}
%% In addition we assume
%% \begin{equation}
%%   \label{eq:xxie0.1}
%% \phi < {4 e^{M + r} \over 4 e^{M + r} + 1}
%% \end{equation}

Rewriting \eqref{eq:xxie0} gives
\begin{eqnarray*}
  u(v) + b u(\delta v) &=& \E u(C) + b \E [u(C); C < \delta v] \\
  &=& \alpha p M^\alpha{1 + b \over 1 - \gamma}\int_{-\infty}^{-M}
  [(1 - \phi) e^r + \phi e^x]^{1 - \gamma} |x|^{-\alpha-1} dx \\
  && + {1 + b \over 1 - \gamma}\int_{-M}^{\delta v}
  [(1 - \phi) e^r + \phi e^x]^{1 - \gamma} d F_u(x) \\
  && + \int_{\delta v}^{\infty}
  [(1 - \phi) e^r + \phi e^x]^{1 - \gamma} d F_u(x) \\
  &=& I_1 + I_2 + I_3
\end{eqnarray*}
Only the first term $I_1$ involves the tail parameters $\alpha$
and $M$ in which we are interested, so we will only compute this
term. The integral of $I_1$ can be written as
\begin{eqnarray*}
&& \int_{-\infty}^{-M}
[(1 - \phi) e^r + \phi e^x]^{1 - \gamma} |x|^{-\alpha-1} dx \\
&=& \int_{M}^{\infty}
[(1 - \phi) e^r + \phi e^{-x}]^{1 - \gamma} x^{-\alpha-1} dx
\end{eqnarray*}
Taylor expansion at $\phi e^{-x} = 0$ gives
\begin{eqnarray*}
  && [(1 - \phi) e^r + \phi e^{-x}]^{1 - \gamma} \\
  &=& ((1 - \phi) e^r)^{1 - \gamma}
  + \sum_{n=1}^{\infty} {1 \over n!}
  {\Gamma(2 - \gamma) \over \Gamma(2 - n -\gamma)}
  ((1 - \phi) e^r)^{1 - n - \gamma} \phi^n e^{-n x}
\end{eqnarray*}
Therefore
\begin{eqnarray*}
&& \int_{-\infty}^{-M}
[(1 - \phi) e^r + \phi e^x]^{1 - \gamma} |x|^{-\alpha-1} dx \\
&=&
{1 \over \alpha} M^{-\alpha} ((1 - \phi) e^r)^{1 - \gamma}
+ \sum_{n=1}^{\infty} {1 \over n!}
{\Gamma(2 - \gamma) \over \Gamma(2 - n -\gamma)}
((1 - \phi) e^r)^{1 - n - \gamma} n^\alpha \Gamma(-\alpha, n M)
\end{eqnarray*}
where $\Gamma(\alpha, x)$ is the upper incomplete gamma
function. For $x > 1$, this function may be expressed as (equation
8.11.2 of {\it NIST Digital Library of Mathematical Functions}
\cite{NIST.DLMF})
\[
\Gamma(a, z) = z^{a - 1} e^{-z}
\sum_{k=0}^\infty (-1)^k {\Gamma(a) \over \Gamma(a - k)} z^{-k}
\]
Thus after re-arranging terms we get
\begin{eqnarray}
  {1 \over 1 + b} I_3
  &=&
  {p \over 1 - \gamma} ((1 - \phi) e^r)^{1 - \gamma} \nonumber \\
  &&
  + \sum_{k=0}^\infty
  {\alpha p ((1 - \phi) e^r)^{1 - \gamma}\over M^{k+1}}
  \prod_{i=0}^{k-1} (\alpha + i)
  \sum_{n=1}^\infty {(-1)^{n-1} \over n! n^{k+1}}
  \left(
  {\phi \over e^{M + r} (1 - \phi)}
  \right)^n \times \nonumber \\
  && \prod_{i=1}^{n-1} (\gamma + i - 1)
  \label{eq:xxie1}
\end{eqnarray}
Observe that the sum over $n$ does not involve $\alpha$. Denote
\[
a_n(k) =
{(-1)^{n-1} \over n! n^{k+1}}
\left(
     {\phi \over e^{M + r} (1 - \phi)}
     \right)^n
     \prod_{i=1}^{n-1} (\gamma + i - 1)
\]
For each given $k$, $a_n(k)$ is an alternating series.
\begin{eqnarray*}
  && {a_n(k) \over |a_{n+1}(k)|} \\
  &=& {
    e^{M + r} (1 - \phi) (n+1)^{k+2}
    \over
    \phi n^{k+1} (n + \kappa - 1)
  } \\
  &=&
  \left(1 + {2 - \kappa \over n + \kappa - 1}\right)
  \left( 1 + {1 \over n} \right)^{k+1}
  {e^{M + r} (1 - \phi) \over \phi}
\end{eqnarray*}

If ${e^{M + r} (1 - \phi) \over \phi} > 1$, it is clear
$a_n(k)/|a_{n+1}(k)| > 1$. It follows that $I_3$ and hence $u(v) +
u(\delta v)$ is monotonically increasing with $\alpha$. If all
investors are guided by and hence seek to maximize the same utility
function, they would all 
have constructed their portfolios by choosing from a set of equities
that have tail indices close to the largest in the market. At market
equilibrium, all actively traded equities would have tail indices that
are indistinguishable from each other.

\textcolor{red}{
  What happens when ${e^{M + r} (1 - \phi) \over \phi} \leq 1$?
}

\subsubsection{very risk-avert utility functions}
In this section we consider power utility functions of
\eqref{eq:power_utility} with $\gamma > 1$. For convenience, let $ĸ =
\gamma - 1$.

@@

\section{Hoga Test}
%% TODO: Show the test statistics of Hoga test applied to a sequence of
%% two concatenated return series. A changes in the scale is characterized
%% by a smooth bump in the test statistics while a change in the tail
%% index is characterized by a jump. Derive a theoretical result to
%% support this?

>>>>

\section{Generalized disappointment aversion}
\label{sec:GDA}
Routledge and Zin (2010) \cite{routledge2010generalized} define the GDA
preferences over discrete states in their formula (2). A slight extension
is to consider a continuum of possible states. Suppose the state variable
$C$ is continuously supported on $(0, \infty)$ and has distribution
function $G(\cdot)$. The utility function of an agent exhibiting GDA
preferences reads
\begin{equation}
  u(v)= \E u(C) - b \int_{0}^{\delta v}
  \left[ u(\delta v) - u(x) \right] dG(x)\label{11}%
\end{equation}
where $\delta > 0$, and $v>0$. Equivalently
\begin{equation}
  \label{eq:xxie0}
  u(v) = \E u(C) - b \E[u(\delta v) - u(C); C < \delta v]
\end{equation}
Here $v$ can be thought of as the certainty payoff equivalent to the
risky payoff $C$. Note that when $b=0$, 
preferences are expected utility; when $\delta=1$ and $b>0$,
preferences follow Gul's (1991) original disappointment aversion.
An agent guided by this utility function will seek to maximize $u(v)$.

\subsection{power-law utility function}
In this section we follow Routledge and Zin
\cite{routledge2010generalized} and assume 
that the utility function of consumption $C$ is a power function%
\begin{equation}
  \label{eq:power_utility}
  u(C)=\frac{1}{1-\gamma}C^{1-\gamma}%  
\end{equation}
where $\gamma > 0$. For simplicity we assume the investor
initially has one unit of wealth. Let $r$ be the return on the
riskless state bond and $X$ be the return on equity. Portfolio weights
are $\left(1-\phi\right)$ and $\phi$ respectively. Then
\begin{equation}
  \label{eq:xxie1}
  %% u(C)=\frac{1}{1-\gamma}V^{1-\gamma} \max \left\{
  %% 0, 1 + (1-\phi) r  +\phi X
  %% \right\}^{1-\gamma} \label{3}%
  C = (1 - \phi) e^r + \phi e^X
\end{equation}

%% \subsubsection[in case gamma > 1]{in case $\gamma > 1$}

\subsubsection[in case gamma > 1]{in case $\gamma > 1$}
In this section we consider power utility functions of
\eqref{eq:power_utility} with $\gamma > 1$. For convenience,
let $\xi = \gamma - 1$. In this case, we assume the following
distribution function $F(\cdot)$ of the equity return:
\begin{equation*}
  F(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x < 0 \\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0
  \end{array}
  \right.
\end{equation*}
Assume $\alpha, \beta > 1$, $0 < p < 1$, $K > 0$. Equation
\eqref{eq:xxie0} can be re-arranged as
\[
u(v) + b u(\delta v))
=
\E u(C) + (1 + b) \E [u(\delta v) - u(C); C < \delta v]
\]
The right-hand-side of the above equation is a functional of the
distribution function of $X$ via eq.\eqref{eq:xxie1}. For convenience
we denote it $\mathcal U(\cdot)$, i.e.
\[
\mathcal U F (\alpha, \beta, K, K', \phi)
= 
\E u(C) + (1 + b) \E [u(\delta v) - u(C); C < \delta v]
\]
\begin{theorem}
  Let
  \[
  G(\alpha, \beta, K, K') =
  \max_{0 < \phi \leq 1} \mathcal U F(\alpha, \beta, K, K', \phi)
  \]
  $G(\alpha, \beta, K, K')$ is monotonically increasing with
  $\alpha$.
\end{theorem}

\begin{proof}
  We show
  \[
  \td{G}{\alpha} > 0
  \]
  Let
  \begin{equation}
    \label{eq:xxie2}
    \hat \phi := \argmax_{0 < \phi \leq 1}
    \mathcal U F(\alpha, \beta, K, K', \phi)
  \end{equation}
  We have
  \[
  G(\alpha, \beta, K, K') = \mathcal U F(\alpha, \beta, K, K', \hat \phi)
  \]
  It follows
  \begin{eqnarray}
    \td{G}{\alpha}
    &=&
    \pd{\mathcal U F}{\alpha}(\alpha, \beta, K, K', \hat \phi)
    + \pd{\mathcal U F}{\phi}(\alpha, \beta, K, K', \hat \phi)
    \pd{\hat \phi}{\alpha}(\alpha, \beta, K, K')
    \label{eq:xxie3}
  \end{eqnarray}
  The definition \eqref{eq:xxie2} implies $\forall \alpha > 1$
  \begin{equation}
    \label{eq:xxie4}
    \pd{\mathcal U F}{\phi}(\alpha, \beta, K, K', \hat \phi) = 0
  \end{equation}
  So the second term of \eqref{eq:xxie3} vanishes. It remains to show
  \[
  \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi) > 0
  \]
  \begin{enumerate}[i.]
  \item When $C < \delta v$ implies $X < -q$ for some $q > 0$, we have
    \begin{eqnarray*}
      \pd{\mathcal U(F)}{\alpha}
      &=&
      - \frac{\partial}{\partial \alpha} \left\{
      {\alpha p K^\alpha \over \xi}
      \left[
      (1 + b)\int_{-\infty}^{-q} +
      \int_{-q}^0
      \right] \right. \\
      && \left.
      \left\{
      \left[
        (1 - \phi) e^r + \phi e^x
      \right]^{-\xi}
      (K - x)^{-\alpha - 1}
      \right\}
      dx \right\} \\
      &=&
      \underbrace{
      {p \over \xi}
      \left[
      (1 + b)\int_{q}^{\infty} +
      \int_{0}^q
      \right]
      \left[
        (1 - \phi) e^r + \phi e^{-x}
      \right]^{-\xi}
      \ln \left(
      1 + {x \over K}
      \right)
            {\alpha K^{\alpha} dx
              \over
              (K + x)^{\alpha + 1}
            }
      }_{L_1}\\
      && - \underbrace{
        {p \over \xi}
        \left[
          (1 + b)\int_{q}^{\infty} +
          \int_{0}^q
          \right]
        \left[
          (1 - \phi) e^r + \phi e^{-x}
          \right]^{-\xi}
        {K^{\alpha} dx \over (K + x)^{\alpha + 1} }
      }_{L_2} \\
    \end{eqnarray*}
    Clearly $L_2 < \infty$. For $L_1$, observe
    \begin{eqnarray*}
      L_1 &<& {
        \alpha K^\alpha p (1 + b)
        \over
        \xi (1 - \phi)^\xi e^{r \xi}
      } \int_0^\infty
      {1 \over (x + K)^{\alpha + 1}}
      \left(1 + {x \over K}\right)
      dx \\
      &=&
      {
        p \alpha (1 + b)
        \over
        \xi (\alpha - 1) (1 - \phi)^\xi e^{r \xi}
      } < \infty
    \end{eqnarray*}
    Let
    \begin{eqnarray*}
      c &:=& {p \over \xi}
      \left[
      (1 + b)\int_{q}^{\infty} +
      \int_{0}^q
      \right]
      \left[
        (1 - \phi) e^r + \phi e^{-x}
      \right]^{-\xi}
      {
        \alpha K^{\alpha} dx \over (K + x)^{\alpha + 2}
      } < \infty
    \end{eqnarray*}
    Define distribution function $H(x): (0, \infty) \to (0, 1)$ by its
    density with respect to the Lebesgue measure
    \[
      {d H(x) \over d x}
      =
      {p \over c \xi}
      \left[
        (1 - \phi) e^r + \phi e^{-x}
        \right]^{-\xi}
      {
        \alpha K^{\alpha}
        \over
        (K + x)^{\alpha + 2}
      }
      \times
      \left\{
      \begin{array}{ll}
        1 & \text{ if } 0 < x < q \\
        (1 + b) & \text{ if } x > q \\
      \end{array}
      \right.
    \]
    Let $Y \sim H$, then
    \begin{eqnarray*}
      \E Y &:=& {p \over c \xi}
      \left[
        (1 + b)\int_{q}^{\infty} +
        \int_{0}^q
        \right]
      \left[
        (1 - \phi) e^r + \phi e^{-x}
        \right]^{-\xi}
      \alpha K^{\alpha} (K + x)^{-\alpha - 1} dx \\
      &=& {\alpha L_2 \over c} \\
      L_1 &=& c \E[Y \ln(Y/K)]
    \end{eqnarray*}
    Notice
    \[
      {
        d^2
        \over
        dx^2
      }
      \left[
        x \ln\left({x \over K}\right)
      \right] = {1 \over x} > 0
      \quad \text{ for } x > 0
    \]
    That is, $x \ln(x/K)$ defined on $(0, \infty)$ is a convex
    function. Now that $L_1, L_2 < \infty$, Jensen's inequality gives
    \[
      {L_1 \over c} = \E [Y \ln(Y/K)]
      >
      \E Y \times \ln(\E Y/K)
      =
      {\alpha L_2  \over c}
      \ln\left(
      {\alpha L_2  \over c K}
      \right)
    \]
  \end{enumerate}
  Therefore
  \[
  L_1 - L_2 > L_2 \left[
    \alpha \ln \left(
    {\alpha L_2 \over c K}
    \right) - 1
  \right]
  \]
  Clearly $L_2 > 0$. So it remains to show
  \[
    \alpha \ln \left(
    {\alpha L_2 \over c K}
    \right) - 1 > 0
  \]
  Observe
  \begin{eqnarray*}
    L_2 &>&
      {p \over \xi}
      \left[
      (1 + b)\int_{q}^{\infty} +
      \int_{0}^q
      \right]
      \left[
        (1 - \phi) e^r + \phi
      \right]^{-\xi}
      {K^{\alpha} dx
        \over
        (K + x)^{\alpha + 1}
      } \\
      &=& {
        p [(K + q)^{-\alpha} b + K^{-\alpha}]
        \over
        \alpha \xi [(1 - \phi) e^r + \phi]^\xi
      } 
  \end{eqnarray*}
\end{proof}

%% \[
%% F(x) = \left\{
%% \begin{array}{ll}
%%     p M^\alpha / |x|^\alpha & \text{ if } x \leq -M \\
%%     F_u(x) & \text{ if } -M < x < M\\
%%     1 - p M^\alpha x^\alpha & \text{ if } x \geq M
%% \end{array}
%% \right.
%% \]
%% where the distribution function $F_u(\cdot)$ satisfies
%% \begin{eqnarray*}
%%   F_u(-M) &=& p \\
%%   F_u(M) &=& 1 - p
%% \end{eqnarray*}
%% In this case we have
%% \begin{eqnarray*}
%%   && u(v) + b u(\delta v) \\
%%   &=&
%%   {
%%     (1 + b) p \alpha M^\alpha
%%     \over
%%     \kappa
%%   }
%%   \int_{-\infty}^{-M}
%%   ((1 - \phi) e^r + \phi e^x)^{-\kappa} |x|^{-\kappa - 1}
%%   dx + \\
%%   &&
%%   (1 + b) \int_{-M}^{\delta v}
%%   ((1 - \phi) e^r + \phi e^x)^{-\kappa} d F_u(x) + \\
%%   &&
%%   {
%%     (1 + b) p \alpha M^\alpha
%%     \over
%%     \kappa
%%   }
%%   \int_{M}^{\infty}
%%   ((1 - \phi) e^r + \phi e^x)^{-\kappa} x^{-\kappa - 1}
%%   dx \\
%%   &=&
%%   I_1 + I_2 + I_3
%% \end{eqnarray*}

@@

%% \section{Hoga Test}
%% TODO: Show the test statistics of Hoga test applied to a sequence of
%% two concatenated return series. A changes in the scale is characterized
%% by a smooth bump in the test statistics while a change in the tail
%% index is characterized by a jump. Derive a theoretical result to
%% support this?
